---
title: 'Reproducible Research: Peer Assessment 1'
author: "Feng"
date: "August 17, 2017"
output:
  word_document: default
  html_document: default
---
This is the first project of the Reproducible Research course of CourseRA's data science specialization. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

This part is to download, unzip and load data into data frame  data .

```{r loadingandprocessing}

library(ggplot2)
library(tidyverse)

        temp <- tempfile()
        download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
        unzip(temp)
        unlink(temp)


data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

This part is to sum steps by day, create Histogram, and calculate mean and median.

```{r steps}
steps_by_day <- aggregate(steps ~ date, data, sum)
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
mean(steps_by_day$steps, na.rm=TRUE)
median(steps_by_day$steps, na.rm=TRUE)
```

Mean is `r mean(steps_by_day$steps, na.rm=TRUE)` and median `r median(steps_by_day$steps, na.rm=TRUE)`.

## What is the average daily activity pattern?

This part is to calculate average steps for each interval for all days, plot the averae number steps per day by interval, and find interval with most average steps.

```{r pattern}
steps_by_interval <- aggregate(steps ~ interval, data, mean)
plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")

 steps_by_interval[which.max(steps_by_interval$steps),1]
```

The 5-minute interval, on average across all the days in the data set, containing the maximum number of steps is `r steps_by_interval[which.max(steps_by_interval$steps),1]`.


## Impute missing values

This part is to impute missing values and compare the imputed data to non-imputed one.

```{r impute}
incomplete <- sum(!complete.cases(data))
imputed_data <- transform(data, steps = ifelse(is.na(data$steps), steps_by_interval$steps[match(data$interval, steps_by_interval$interval)], data$steps))
imputed_data[as.character(imputed_data$date) == "2012-10-01", 1] <- 0
steps_by_day_i <- aggregate(steps ~ date, imputed_data, sum)
hist(steps_by_day_i$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")

hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="red", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("blue", "red"), lwd=10)
mean(steps_by_day_i$steps)
median(steps_by_day_i$steps)
mean(steps_by_day_i$steps) - mean(steps_by_day$steps, na.rm=TRUE)
median(steps_by_day_i$steps) - median(steps_by_day$steps, na.rm=TRUE)
sum(steps_by_day_i$steps) - sum(steps_by_day$steps)

```

The imputed data mean is `r mean(steps_by_day_i$steps)` and median is `r median(steps_by_day_i$steps)`. The difference between the imputed mean and non-imputed mean is `r mean(steps_by_day_i$steps) - mean(steps_by_day$steps, na.rm=TRUE)`.The difference between the imputed median and non-imputed median is `r median(steps_by_day_i$steps) - median(steps_by_day$steps, na.rm=TRUE)`.The difference between total number of steps between imputed and non-imputed data is `r sum(steps_by_day_i$steps) - sum(steps_by_day$steps)`. 


## Are there differences in activity patterns between weekdays and weekends?

This part is to create a plot to compare and contrast number of steps between the week and weekend.

```{r activity}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$date)),weekdays), "Weekday", "Weekend"))

steps_by_interval_i <- aggregate(steps ~ interval + dow, imputed_data, mean)

library(lattice)
xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```
